Recommentations: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations

### 2024-09-08 Corrections_SAM_SS3_Krill_481.Rmd
wp2

---
title: "Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
subtitle: "WG-FSA 2024. Working Paper"
author:
  - Mauricio Mardones^[Universidad de Magallanes, mamardon@umag.cl, Milenio Base]
  - Lucas Krüger ^[Instituto Antártico Chileno, Milenio Base]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Milenio Base]
  - César Cárdenas^[Instituto Antártico Chileno, Milenio Base]
  - George Watters^[Ecosystem Science Division, Southwest Fisheries Science Center, NOAA, USA]
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

```{r setup1, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(eval=TRUE,
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=50), 
                      tidy=TRUE)
```

```{r message=FALSE, eval=TRUE}
library(here)
library(kableExtra)
library(ggpubr)
library(tibble)
library(readxl)
library(openxlsx)
library(r4ss)
dir1<-here("s1")
```

# BACKGROUND

This study addresses key recommendations highlighted in the preliminary report of the Working Group on Statistics, Assessment, and Modelling (WG-SAM-2024/26) [@Mardones2023] regarding the integration of diverse information sources for understanding the population dynamics of Antarctic krill (*Euphausia superba*) within an integrated stock assessment model, particularly in the context of environmental change and fisheries management challenges. The main observations were: 1) Concerns about using a length-based model due to the scarcity of direct age data, as this approach avoids the approximations required when converting length data to age data; 2) The need to consider the impact of changes in trawl designs over time and to reassess the assumption that the Antarctic Peninsula is a closed system; 3) The limitations of using krill length and biological composition data from fishing vessels, given the significant variability and potential biases, and the importance of understanding the interaction between fisheries and krill-dependent predators; 4) A recommendation to present a progression of model implementations from simple to complex to better evaluate model assumptions, investigate patterns in residuals, and assess the realism of parameter values used, while also integrating recent findings on krill stock hypotheses, with special attention to harvest rate and management procedures.Each observation will be explained in theoretical terms and, where possible, analyzed in detail. This approach reflects an ongoing work in progress that is continuously refined based on interactions with the CCAMLR Working Groups. These interactions provide valuable feedback that helps to shape and improve the study's methodology and findings. It should be noted, of this approach is vital for contributing to management strategies based on quantitative methods that provide indicators of krill health, integrating ecosystem aspects into the assessment.

@@

# METHODOLOGY

The study area for the assessment of krill populations in the Antarctic Peninsula is Subarea 48.1 and strata subdivision to account for spatial heterogeneity [@Dornam2021] (Figure \@ref(fig:mapa)). The Antarctic Peninsula region is vast and encompasses a variety of habitats and environmental conditions that can influence the distribution and abundance of krill. Therefore, dividing the study area into strata allows to compartmentalize information related to fishing and surveys to model krill populations by considering local variations in structure population.

## Data Sources

Multiple sources of data to inform the population dynamics model are included in this analysis. These data sources include: 1) Data from the monitoring program of the krill fishery were used, as CPUE and length compositions, which have been systematically collected on board fishing vessels by scientific observers as part of the CCAMLR Scheme of International Scientific Observation (SISO) program. This database has information about vessel, nationality, georeferenced, among others; 2) Data collected from scientific AMLR research cruises and monitoring programs specifically designed to assess krill populations, including: data on abundance by strata, and biological characteristics of krill collected through net sampling. All this data was processed in @Dornam2021; and length compositions from AMLR survey (provided by Ecosystem Science Division, NOAA) was handled according to the evaluation template requirements (Suppl. Mat. 2); 3) For this analysis, chlorophyll-a concentration (Chl) was assessed as a primary environmental driver. Chl data was sourced from Bio-Geo-Chemical, L4 (monthly and interpolated) satellite observations, obtained through the E.U. Copernicus Marine Service Information (doi.org/10.48670/moi-00021; doi.org/10.48670/moi-00148). This data provides a detailed understanding of primary productivity levels, which are critical for evaluating ecosystem health and  krill dynamics in the WAP. We fitted a GLMM to predict Length with Year (formula: Length ~ Year + Chl + Sea Ice Concentration (SIC) + Sea Surface Temperature (SST)). The model included spatial component (cell) as random effect (formula: ~1 | cell). Krill length compositions from fishery data was correlated with environmental variables by using Pearson correlations test, and General Linear Mixed Model (GLMM) [@Bates2015; @Makowski2020] (Suppl. Mat 3); 4) Predator abundance data from "MAPPPD". Although the penguin colonies that live within the study area behave differently in terms of dynamics throughout the analyzed series, we systematized a vectorized indicator for the purposes of considering it as a fleet to add like M2 in an integrated stock assessment.Data handling for modeling purposes was elaborated with *tidyverse* library  and dependences [@Wickham2019].

## Conceptual Model

A conceptual model for krill in SubArea 48.1 includes the five spatial strata described previously. Despite differences among strata, the krill population is closed for emigration and immigration, and their spatial distribution and interactions occur within this single, cohesive population (Figure \@ref(fig:concem)).Ecological and environmental variables that have an impact on the krill population include predator abundance (e.g., whales, seals, penguins), phytoplankton concentration, and nutrient levels. The model considers interannual fluctuations of growth dynamics , maturity and reproduction, with a spawning pulse per year.

## Statistical model Stock Synthesis (SS3)

Like most invertebrates, krill is a species for which aging is complex [@Punt2013]. Therefore, a size-structured model with age dynamics is used, which is transformed through a transition matrix based on the probabilities of age groups relative to the detailed structure. Integrated models have the capacity to reproduce the age dynamics of these populations and simultaneously transform them into population variables. The stock assessment model to krill was configured using Stock Synthesis (SS3 hereafter) [@Methot2013] with the updated version 3.30.21. SS3 is a widely used software tool for assessing fish and invertebrate populations widely used in the main regional fishing organizations to have approximations of the population dynamics of the exploited resource as well as to support management decisions. SS3 is a structured stock evaluation model, in the class of models called *"Integrated stock evaluation analysis model"* and has a set of sub-model that simulates growth, maturity, fecundity, recruitment,  and mortality processes, and observation, with expected values for different types of data. The model is coded in Automatic Differentiation Model Builder (ADMB) [@Fournier2012; @Methot2013: @methot2020stock] with estimation parameters. The methodology employed by Stock Synthesis involves a comprehensive and integrated approach, utilizing various data sources and modeling techniques to estimate the main population variables of krill in WAP, and unlike other integrated models, it has features that allow the integration of information related to the ecosystem, such as predators and environmental variables that have proven influence on the analyzed population. All analysis were executed in  R-CRAN [@R-base]and the graphical interface of the SS3 through *r4ss* [@Taylor2019]  and *ss3diags* packages [@Winker2023]. Life history parameters, like growth, weight-length relation, natural mortality and maturity was used as priors to model initial condition was taken from @Smith2023a, @Maschette2020 and @Kinzey2011 and can be found in Table \@ref(tab:parainit). Source data and temporal scale used in this stock assessment is showed in Figure \@ref(fig:datas2).

### Environmental variable modelling in `SS3`

One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of krill population dynamics and SS3 addresses these approaches [@methot2020stock]. The estimated level of krill recruitment depends on the spawning biomass from the previous season, an environmental time series—specifically, the 2000-2022 Chlorophyll-a time series—and a log-bias adjustment.

\[
E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
\]

\(R\) represents the variability of deviations, adding to the variance caused by environmental factors. \(SpBio\) represent spawning biomass, \(envdata\) is vector of time series of environmental variable. Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual \(R\) should be reduced. However, the model does not automatically adjust for this.

Furthermore, this integrated model took into account environmental variables, as chlorophyll (Chl), known to influence krill abundance and reproductive outputs. These environmental factors play a critical role in shaping krill habitat suitability and productivity, thereby affecting their population dynamics. By integrating environmental data into our model, we aim to elucidate the relationships between environmental conditions and krill abundance, enabling more robust predictions of krill population dynamics under future climate scenarios.

### Predator components modelling in `SS3`

This stock assessment approach consider predator-prey interactions as a key driver of krill population dynamics. By incorporating information on predator  abundance and feeding ecology, our model provide a more comprehensive understanding of the tropic interactions shaping krill dynamics.  In this analysis we incorporated penguin population index as predator *"fleets"* as we describe previously. This overall natural morality (M) to include explicit mortality (M2) caused by each major predator as M = M1 + sum(M2) [@methot2020stock], and in this way transfer to total mortality in form Z = M1 + M2+ F.

All this component and path to construct models, diagnostics and check outputs is represented in Figure \@ref(fig:path).

## Stock Assessment Scenarios

To consider the uncertainty associated with the krill nature scenarios, we considered a set of configurations associated with spatial heterogeneity, influence of life history parameters on the estimates and uncertainty associated with the relationship that exists between spawning biomass and krill recruitment. The methodology involves comparing the statistical performance of the model regarding these configurations. We have eight scenarios that were used to test different options in modeling about main considerations in assessment of krill population, and for comparative purposes, we consider `s1` as the base model (Table 1).

| Scenario | Description                                                                          |
|:------------:|:---------------------------------------------------------|
|   s01    | Fishery and Survey (AMLR) data, Predator, Environmental data, aggregated at 48.1 level               |
|    s1    | Fishery and Survey (AMLR) Length data, CPUE, Catch by strata. Predator and Env data |
|    s2    | "s1" without Stock-Recruit relation                                      |
|    s3    | "s1" Beverton & Holt S-R relation weak (0.9 steepness)                   |
|    s4    | "s1" BH S-R relation strong (0.6 steepness)   |
|    s5    | "s1" BH S-R relation mid-strong estimated    |
|    s6    | "s1" Ricker S-R relation estimated    |
|    s7    | "s1" w/ set of parameters estimated in @Mardones2024par   |

: Scenarios to modelling dynamics in krill

For the purposes of this working document, a comparative analysis of models `s01`, `s1` and `s7` will be evaluated, given that they contain elements such as aggregate data, spatially structured data and another set of life history parameters according to @Mardones2024par (unpublished).

# RESULTS

## Diagnosis Base Model

All models tested was the final gradient using maximum likelihood is achieve and was relatively small (< 1.00e-04), and the Hessian matrix for the parameter estimates was positive, which shows that the performance is within what was expected for this type of analysis. About goodness-of-fit, Figure \@ref(fig:res1) and Figure \@ref(fig:res1a) display the residuals of length of krill captured by fleets over the years in `s1`. In the first graph, we observe multiple fishing fleets (FISHERYBS, FISHERYEI, FISHERYGS, FISHERYJOIN, FISHERYSSIW) and a survey (SURVEYBS). The second graph shows surveys (SURVEYEI, SURVEYGS, SURVEYJOIN) and predator data (PREDATOR) for `s1`. For many fleets and surveys, greater consistency in krill sizes is observed after the year 2000. However, there was variability in the width of the distributions over time, which may reflect changes in abundance or sampling selectivity. For instance, in FISHERYBS and FISHERYEI, the distributions widen and slightly shift towards larger sizes in recent years, indicating a possible change in the krill population or fishing practices. The residuals from the model fit suggest that the model captures the inter annual and inter-fleet lengths variability well, although some deviations indicate that additional unmodeled factors might be influencing the observed sizes (Figure \@ref(fig:resall),  \@ref(fig:fitcom)).

## Retrospective analysis

Retrospective analysis shows the pattern of bias that exists in the models and is one of the ways in which we have identified that in all the tested scenarios there is a pattern, however, model S1 is the one that performs better. @Carvalho2021b indicate that values of rho parameter that fall outside (-0.15 to 0.20) for SSB for longer-lived species, or outside (-0.22 to 0.30) for shorter-lived species indicates an undesirable retrospective pattern (Figure \@(fig:retros1)). To evaluate the overall model fit of the relative abundance indices and composition data in `s1`, we used the joint-index residual. Overall, `s1` had a good performance when evaluating the lengths compositions (14.2%), however  RMSE showed a low predictive power with respect to the indices (71.3%). A loess-smoother indicated there appeared to be increased variability in the residuals of model fit to CPUE over time (Figure \@ref(fig:rmses1)).

## Likelihood profile in spatial Model `s1`

For `s1` scenario, the gradient of the likelihood profile for the penalty on the index deviations was greater than other data sources. The second strongest gradient in the log-likelihood profile was observed for the length compositions Figure \@ref(fig:likepro). The gradient of the likelihood profile supported by the length-composition data is higher than those supported by the penalty for the recruitment deviates and CPUE indices.

Once the performance of the model was corroborated, it was possible to have an estimate and forecasting of the main population variables over the years in absolute terms, which can be seen in the Table \@ref(tab:mainvar)

# RECOMMENDATIONS SAM - 23

The main recommendations were extracted from @CCAMLRWGSAM2024

*2.3 Noting that the authors recalled a previous independent review of an age-based integrated stock assessment for Antarctic krill which encouraged the development of such assessments (Thomson, 2016), the Working Group highlighted that the other reviewer on that panel noted that a length-based model could be considered due to the sparsity of direct age data (de Lestang, 2016). It also noted that this would avoid the approximations required when converting length data to age data. The Working Group further recalled that a similar comment had been made (WG-SAM-2023, paragraph 4.3) in relation to a pilot Casal2 age-based assessment (WG-SAM-2023/25). *

@@

We consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data. We recognize that using a length-based model could reduce the need for approximations when converting length data to age data. We plan to explore this alternative by developing a length-based model to compare its performance and results with the current age-based model. Additionally, we will incorporate standard model diagnostics, as suggested, similar to those used in toothfish assessments, to ensure transparency and facilitate understanding of model performance.

Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017]. In this sense, when reading of hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap. Length offers insights into the age structure of the population, as there is a correlation between age and length. However, it's still necessary to convert catch-at-length data into catch-at-age data. Several methods are available to achieve this [@ICCA2003].

#### Age-Length Keys (ALKs)

Put simple, age-length keys (ALKs) are generated through the ageing of a sub-sample of the population, and used to convert larger length samples from a population into ages. ALKs describe distributions of size for each age, and the relative number of individuals at each age, i.e. they represent a matrix detailing the probability that a fish of a given length is of a particular age. Once such a key is available, samples of fish that were only measured for length can be distributed over age groups according to the key. The use of ALKs assumes that the sample of aged fish and the sample of fish measured for length are simple random samples from the same population. Then, the probability that a fish is of a particular age, given its length is the same for both samples.
The ALK should generally be applied to length data from the same time period, since variability in recruitment and survivorship at age will change the age-length composition over time, and hence the number of survivors at age used to weight the size at age compositions will vary. The ALK may need to be seasonal if growth is temporally distinctive or seasonal migration occurs. A single ALK should only be applied to size data from a number of years if growth is reasonably stationary, and the approach of Kimura and Chikuni (1987) is used. Suitable justification should be presented for multi-seasonal or multi-annual application of ALKs. It must also be noted that the application of an ALK derived from a single time of year can cause serious bias if used to compute catch at age for the entire year.
Age-at-length data from hard parts should be combined into suitable length groups for required gears, time periods and locations. The size of these length groups will depend on the spread of lengths found within the catch length frequencies, the growth rate of the species, and the variability in length-at-age. Length-stratified sampling is required to ensure that the required number of fish (determined by the variability of length-at-age) is available over all length groups. Developing ALKs represent is laborious, and hence the optimum collection of information is desirable. Formulae exist to estimate the number of age determinations and length measurements necessary to guarantee a given level of accuracy. Oeberst (2000) developed a universal cost function for ALKs, for example [@ICCA2003].

The proportion at age is calculated as:
$$
\frac{\text{Number at age for a length group}}{\text{Number of fish aged in that length group}}
$$

@@

If the ALK does not contain data for all the length groups in the length distribution then data in the length distribution may be assigned to adjacent length groups where data are present in the ALK. An appropriate age- length distribution by gear group and time period is then produced. Care must be taken since considerable biases can result, particularly at large lengths where individuals may be distributed over a wide range of ages.

The numbers at age by length group are summed over the length range to give numbers at age. The variances are also summed over the length groups and the two components labelled variance due to ageing and variance due to length sampling. This gives the age composition for the required time period. Numbers at age for all gears can be calculated as:

@@

The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK. The method can work well where the distributions of size for each age are close to those in the length frequency, but convergence can be slow. The reader is referred to the paper for more information.

Then Numbers-at-age in the model were converted to numbers-at-length (mm) using a von Bertalanffy growth relationship withestimated parameters L~inf~, k, and v(Eq. (A.6)). This length-at-age relationship, mediated by survey catchability and selectivity (Eqs.(A.2) and (A.3)), was used to compare survey biomasses with model estimates of vulnerable biomass (Eq. (A.4)). Weight-at-length was assumed known and calculated using Eq. 3 from Hewitt et al. (2004)(W(L) = 2.236 × 10-6L3.314). (REF Kindzey)

Another way that we implemented in `SS3` is regarding *platoons*. This approach allows for the exploration of survivorship that varies with size. A value of 1 will not create additional groups. Odd-numbered values (e.g., 3, 5) will divide the overall morph into that many groups, resulting in smaller, larger, and average growth groups. Increasing the number of groups slows down the model's performance, so it's recommended not to exceed a value of 5. The fraction of each morph assigned to each group can be either user-defined or approximated using a normal distribution. When multiple groups are designated, an additional input is required to specify the ratio of variability in size-at-age between groups versus within groups. This ratio is used to distribute the total growth variability. The size-at-age for each group is then calculated as a factor (based on the between-group and within-group variability) times the size-at-age of the central morph, which is determined from the growth parameters specific to that Growth Pattern x Sex combination [@methot2020stock].

@@

In `SS3` colud be conditioned in this way;

@@

Regarding to observation in SAM -2023 (paragraph 4.3) that say;

@@

@@

*2.4 While noting that this study constituted interesting and important work, the Working Group noted that some issues needed further consideration including the change of trawl designs over the course of the collection of the data used as inputs to the model (e.g. changes in mesh sizes and the presence of fine-mesh codend inserts) and the likely invalid assumption of the Peninsula as being a closed system. It further highlighted the need to discuss the development of a standardized data collection plan to support the ongoing revision of the krill fishery management approach.*

We appreciate the identification of potential issues related to changes in trawl designs and the assumption of the Peninsula as a closed system. We will review the trawl data to assess the impact of changes in mesh sizes and fine-mesh codend inserts on the model inputs. Moreover, we acknowledge the need to reconsider the assumption of the Peninsula being a closed system, and we will explore alternative approaches to address this. Moving forward, we will engage in discussions to develop a standardized data collection plan that aligns with the revision of the krill fishery management approach, ensuring that future data inputs are consistent and reliable.

@@

- **Response**: We understand the concerns raised regarding the suitability of krill length and biological composition data from fishing vessels for this modeling approach. We agree that significant differences in length composition between commercial and scientific trawls, as well as among different vessels, pose challenges. To address this, we will review the selection criteria for the data used in our model and consider incorporating data exclusively from standardized scientific trawls to improve the reliability of the model outputs. We also acknowledge the importance of understanding the interaction between fisheries and krill-dependent predators. We will explore opportunities to integrate regular observations and data on the spatial overlap of fishing zones and predator foraging areas into the model, possibly leveraging existing ecosystem observations like those from RV Atlantida.

@@

 - **Response**: We agree that a progression from simple to more complex model implementations would facilitate understanding and provide a clearer evaluation of the model assumptions. We will outline a plan to present this progression in future work, beginning with a simpler model and gradually incorporating complexity. Furthermore, we acknowledge the presence of patterns in the residuals and the need to investigate these further. We will conduct a thorough analysis of the residuals to identify any underlying issues and reassess the realism of the parameter values used in the model. In addition, we will incorporate recent findings from SKEG regarding the krill stock hypothesis to ensure that our model assumptions are up-to-date and reflect the latest scientific understanding.

*show  Harvest Rate*

*Diagram management procedure*

>>>>

---
title: "Recommentations: Integrated approach to modeling krill population dynamics in Western Antarctic Peninsula. Spatial and ecosystem considerations"
subtitle: "Appendix WP- SAM 2024/26"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.3
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---
\newpage
\tableofcontents
\newpage

```{r setup1, echo=FALSE}
rm(list = ls())
knitr::opts_chunk$set(eval=TRUE,
                      echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.align = 'center',
                      fig.pos = "H",
                      dev = 'jpeg',
                      dpi = 300,
                      tidy.opts=list(width.cutoff=80), 
                      tidy=TRUE)
```

```{r message=FALSE, eval=TRUE}
library(here)
library(kableExtra)
library(ggpubr)
library(tibble)
library(readxl)
library(openxlsx)
library(r4ss)
library(flextable)
dir1<-here("s1")
```

@@

# OBSERVATION SAM - 23

The main recommendations were extracted from @CCAMLRWGSAM2024-

## Observation 1

*2.3 Noting that the authors recalled a previous independent review of an age-based integrated stock assessment for Antarctic krill which encouraged the development of such assessments (Thomson, 2016), the Working Group highlighted that the other reviewer on that panel noted that a length-based model could be considered due to the sparsity of direct age data (de Lestang, 2016). It also noted that this would avoid the approximations required when converting length data to age data. The Working Group further recalled that a similar comment had been made (WG-SAM-2023, paragraph 4.3) in relation to a pilot Casal2 age-based assessment (WG-SAM-2023/25).*

@@

We consider that a length-based model as an alternative to the age-based approach, especially given the scarcity of direct age data in krill monitoring. Aging techniques for krill have not been available historically, although some recent progress has been made [@Kilada2017]. In this sense, when reading of hard parts is difficult like in krill, catch-at-length data are more plentiful, as the collection of length information is relatively cheap. Length offers insights into the age structure of the population, as there is a correlation between age and length.  So far, all integrated models for krill have been of the Catch-at-Length type [@Kinzey2015a; @Kinzey2019a; @Wang2021]. This means they are approaches that internally model age through growth parameters and the Age-length-key (ALK). To perform this conversion, various methods have been used, but the AKL matrix is perhaps the most important one [@ICCA2003].

In summary, ALK are generated by aging a sub-sample of a population and used to estimate the age distribution from larger length samples. ALKs assume both aged and measured fish are random samples from the same population and should be applied within the same time period to avoid biases. Seasonal or multi-annual applications require careful justification, as using an ALK from a single period can introduce significant bias. Length-stratified sampling is essential, and the process of developing ALKs is labor-intensive, requiring optimal data collection to ensure accuracy  [@ICCA2003]. Formulae exist to estimate the number of age determinations and length measurements is calculated as;

$$
\frac{\text{Number at age for a length group}}{\text{Number of fish aged in that length group}}
$$

@@

If the ALK lacks data for certain length groups, data from those groups may be assigned to adjacent ones in the ALK. This process generates an age-length distribution by gear and time period, but caution is needed to avoid significant biases, especially for larger lengths with a broad age range. The numbers at age are summed across length groups, combining variances from ageing and length sampling to produce the age composition for the specified period. Numbers at age for all gears can then be calculated. Numbers at age for all gears can be calculated as:

@@

The proportions-at-age are then adjusted to find the best fit between the observed size frequency data and that predicted by the proportion-at-age and the ALK. The majority of ALKs are based on growth parameters according to the most suitable formula for the species. In the case of krill, these parameters are derived from the von Bertalanffy growth curve. In @Mardones2023, using `SS3` stock assessment catch-at-lenght model, we use a von Bertalanffy growth relationship with estimated parameters L~inf~, k, and CV. This length-at-age relationship, mediated by survey catchability and selectivity, was used to compare survey biomasses with model estimates of vulnerable biomass (Eq. (A.4)). Weight-at-length was assumed known and calculated using @Maschette2020 parametres.

```{r AKL}
alk_matrix <- base.model1$ALK[,,1] # Ajusta las dimensiones según tu matriz

# Convertir la matriz en un data.frame
alk_df <- as.data.frame(as.table(alk_matrix))

# Renombrar columnas para mayor claridad
colnames(alk_df) <- c("Length", "TrueAge", "Value")

# Asegúrate de que las columnas Length y TrueAge sean numéricas
alk_df$Length <- as.numeric(as.character(alk_df$Length))
alk_df$TrueAge <- as.numeric(as.character(alk_df$TrueAge))

# Asegúrate de que la columna Value sea numérica
alk_df$Value <- as.numeric(alk_df$Value)

# Crear el plot usando ggplot
ggplot(alk_df, aes(x = TrueAge, y = Length, fill = Value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(x = "True Age",
       y = "Length",
       fill = "Value") +
  theme_bw()+
  theme(legend.position = "none")
```

One way to avoid dependence on converting variables such as size selectivity to age implemented in `SS3` is through the use of *platoons*. This approach allows for the exploration of survivorship that varies with size against the age. A value of 1 will not create additional groups. Odd-numbered values (e.g., 3, 5) will divide the overall morph into that many groups, resulting in smaller, larger, and average growth groups. Increasing the number of groups slows down the model's performance, so it's recommended not to exceed a value of 5. The fraction of each morph assigned to each group can be either user-defined or approximated using a normal distribution. When multiple groups are designated, an additional input is required to specify the ratio of variability in size-at-age between groups versus within groups. This ratio is used to distribute the total growth variability. The size-at-age for each group is then calculated as a factor (based on the between-group and within-group variability) times the size-at-age of the central morph, which is determined from the growth parameters specific to that Growth Pattern x Sex combination [@methot2020stock].

@@

In `SS3`, this approach was conditioned as follows;

@@

in this way, regarding to observation in SAM -2023 (paragraph 4.3);

@@

Since we don't have age data monitoring program, we plan to explore *platoon* alternative to compare its performance and results with the current length-based model using currently in WG-SAM-2024/26.

## Observation 2

@@

Regarding @WGSAM2114, they provide diagnostic plots for the 2021 assessment model for Antarctic toothfish (*Dissostichus mawsoni*) in the Ross Sea region, following the recommendations of WG-SAM-2015 (SC-CAMLR-XXXIV 2015 Annex 5), and for which it is recommended to follow the diagnostic structure for this analysis. @WGSAM2114 has several way to test the perfomance of stock assessment in toothfish, among them are;  progrssion catch-at-age, likelihood profile, Maximum posterior density (MPD), Autocorrelation plots (ACF), Pearson residuals by sex, Markov Chain Mote Carlo priors and posteriors distributions and fits to ages compositions.

Some of the analyses presented in @WGSAM2114 are also implemented in @Mardones2023.

- 1. Likelihood profile

```{r likepro1, out.width='70%', fig.cap="Log-likelihood profiles for R0 for the various data components included in Antarctic krill"}
knitr::include_graphics('Figs/Profile_s1.png')
```

- 2. Pearson residuals to fishery and surveys

```{r resfleet, out.width='50%', fig.cap="Pearson residuals to fishery. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected). in Antarctic krill"}
knitr::include_graphics('s1/plots/comp_lenfit_residsflt6mkt0.png')
```

```{r ressurv, out.width='50%', fig.cap="Pearson residuals to survey. Closed bubbles are positive residuals (observed > expected) and open bubbles are negative residuals (observed < expected). in Antarctic krill"}
knitr::include_graphics('s1/plots/comp_lenfit_residsflt7mkt0.png')
```
- 3. Numbers by length

```{r numsex, out.width='50%', fig.cap="Beginning of year expected numbers at length in (max ~ 2890.5 billion) in Antarctic krill"}
knitr::include_graphics('s1/plots/numbers6_len_sex1.png')
```
- 4. Monte Carlo Prior distributions

```{r marcov, out.width='90%', fig.cap="Each parameter distribution plots (plot 1 of 8). Deviation parameters are not included. Plotting range is equal to input limits on parameters in Antarctic krill"}
knitr::include_graphics(c('s1/plots/parameter_distributions_page1.png',
                          's1/plots/parameter_distributions_page2.png',
                          's1/plots/parameter_distributions_page3.png',
                          's1/plots/parameter_distributions_page4.png',
                          's1/plots/parameter_distributions_page5.png',
                          's1/plots/parameter_distributions_page6.png'))
```

- 5. Autocorrelation recruitment

- 6. Likelihood values

It is important to note that there is a degree of consistency in how the performance of stock assessment models for toothfish and krill is evaluated. For krill, the primary diagnostics conducted adhere to the "best practices" outlined in @Carvalho2021b work. This recipe of steps for testing model performance is illustrated in Figure \@ref(fig:recipediags).

```{r recipediags, out.width='80%', fig.cap="Conceptual process flow chart illustrating a series of interconnected diagnostic tests recommended when developing a base model."}
knitr::include_graphics('Figs/recipediags.png')
```

While we acknowledge the work done by @WGSAM2114 in toothfish stock assessment, we propose a converge towards these standard practices, regardless of the assessment platform and the structure of the integrated model.

## Observation 3.1

*2.4 While noting that this study constituted interesting and important work, the Working Group noted that some issues needed further consideration including the change of trawl designs over the course of the collection of the data used as inputs to the model (e.g. changes in mesh sizes and the presence of fine-mesh codend inserts)...*

This observation raise implications for fishery selectivity. To develop this we configured  a selectivity block in an alternative scenario (`s9`) and compared its performance with the base model (`s1`). This is configured for all strata in 48.1. Tn this case, selectivity blocks are configured in `SS3` as show this template lines in `control` file:

```{r eval=FALSE, echo=TRUE}
1 #_Nblock_Patterns
2 #_blocks_per_pattern
#_begin and end years of blocks
c(1980,2006,2006,2020)
```

The comparison of the main evaluation results with both scenarios is presented in the estimates of the population variables of biomass (Figure \@ref(fig:comparebio)) and recruitment (Figure \@ref(fig:comparerec)). It can be observed that there are no significant differences, at least in the state variables.

```{r comparebio, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
knitr::include_graphics('Figs/compare4_Bratio_uncertainty.png')
```

```{r comparerec, out.width='80%', fig.cap="biomass estimates  in s1 and s9 scenarios."}
knitr::include_graphics('Figs/compare10_recruits_uncertainty.png')
```
## Observation 3.2

*...and the likely invalid assumption of the Peninsula as being a closed system. It further highlighted the need to discuss the development of a standardized data collection plan to support the ongoing revision of the krill fishery management approach.*

Regarding stock unit issue, in the last online workshop of the SCAR Krill Expert Group (SKEG), conclude that The opinion of the SKEG was that the southern boundary of the ACC (sbACC) serves as an
important potential pathway for krill flux into Subarea 48.1, from the Bellingshausen Sea to the
northern Antarctic Peninsula, and into Subarea 48.2. The Weddell Front and Antarctic Slope
Front were also considered important pathways for krill flux into Subarea 48.2, but with focus in 48.1, with a importance of retention and flux (source to destination relation) identified by SKEG
should be taken into account when allocation of catch limits, particularly the importance
of retention in the Gerlach Strait, Bransfield Strait and Joinville Island strata, and the
directions of the important fluxes [WGEMM2439] (Figure \@ref(fig:ksh)).

There are two important issues related to the current stock assessment approach presented in WG-SAM 24/27. The first one, while we understand that there are krill flows to and from Area 48.1, the challenge of having time series data to implement a model that covers most of Subarea 48 becomes difficult. However, we note that as monitoring programs provide the necessary data to implement models of this nature, we are open to a more comprehensive approach in the Southern Ocean. Secondly, the current model implemented in `SS3` considered the spatial heterogeneity of the krill stock structure within Area 48.1, as illustrated in the Figure \@ref(fig:ksh), however, a model that involves movement between strata has not yet been developed, and is part of the ongoing work presented at the SAM.

```{r ksh, out.width='80%', fig.cap="Graphic representation of Krill flux and retention in management strata proposed in the revised krill fishery management strategy within Subarea 48.1, as indicated by the SKEG WG"}
knitr::include_graphics('Figs/KSH.png')
```
\newpage

## Observation 4

@@

We appreciate the identification of potential issues related to changes in trawl designs and the assumption of the Peninsula as a closed system. We will review the trawl data to assess the impact of changes in mesh sizes and fine-mesh codend inserts on the model inputs.  According to the `C1` form from the database provided by the secretary through the SISO program, it is possible to identify (Figure \@ref(fig:trawl)) that there was a change in technique starting in 2006.

```{r trawl, out.width='70%', fig.cap="Display of leght compositios with different trawl technique use in 48.1 subarea."}
knitr::include_graphics('Figs/trawl.png')
```

This consideration was implemented as a new scenario (`s9`) as previously described.

## Observation 5

@@

We agree that a progression from simple to more complex model implementations would facilitate understanding and provide a clearer evaluation of the model assumptions. However, and regarding this observation, a visual inspection of Pearson residuals can lead to a mistaken conclusion about the model *goodness-of-fit*. We have integrated a quantitative element to evaluate this type of test. The plotting options are kept mainly to those provided by `r4ss`. With `SSplotRunstest()`, runs tests results are illustrated for mean lengths of size composition data from the  `SS3`model. Green shading indicates
no evidence (p > 0.05) and red shading evidence (p < 0.05) to reject the hypothesis of a randomly distributed time-series of residuals, respectively. The shaded (green/red) area spans three residual standard deviations to either side from zero, and the red points outside of the shading violate the ‘three-sigma limit’ for that series. We plot this in Figure \@ref(fig:runtest).

```{r runtest, out.width='40%', fig.show='hold',  fig.cap="Runs test plot and Joint residual plot for mean lengths from fits length composition data from different fleets in Antarctic krill."}
sspar(mfrow = c(5, 2), plot.cex = 0.8)
knitr::include_graphics(c('Figs/run1.png',
                          'Figs/run2.png',
                          'Figs/run3.png',
                          'Figs/run4.png',
                          'Figs/run5.png',
                          'Figs/run6.png',
                          'Figs/run7.png',
                          'Figs/run8.png',
                          'Figs/run9.png'))
```

To facilitate automated processing, results from several diagnostic tests can also be called as table. Table \@ref(tab:runtable)
